%!TEX root = ../Thesis.tex
\chapter{Synopsis}
\label{sec:synopsis}
The following chapters are comprised of four journal papers that are supplemented with two conference papers and four workshop papers, of which all are peer-reviewed or submitted to peer-reviewed journals. I combine several papers into topical chapters for conciseness. This thesis follows a data science workflow, starting with exploratory data analysis to gain insight to the specific geology and 4D seismic. I go on to present groundwork on machine learning and data processing on 4D seismic data of the Danish North Sea. Based on this groundwork, I developed a method for 4D seismic inversion and a novel unsupervised 3D time-shift extraction method for 4D seismic. This chapter summarizes these papers and places them in the appropriate context for the thesis.

\input{chapters/03-Synopsis/01-dataprep.tex}

\input{chapters/03-Synopsis/02-foundations.tex}

\input{chapters/03-Synopsis/03-inversion.tex}

\input{chapters/03-Synopsis/05-timeshifts.tex}

\section{Contributions of this Study}

This thesis contributes tangible \acl{ml} applications in geoscience that solve real-world problems as well as work that contributes to the fundamental understanding of signal processing and \aclp{nn} in non-stationary physics.

The explorational work of this thesis validates the impact of image processing for enhancing the resolution of seismic data and automatic fault extraction. This work investigated the scale-gap between local \acl{bhi} and regional seismic data \citep{aabo2017correlation, aabo2018integrated}. The exploration of \acl{bsem} data introduced a novel unsupervised method to extract chalk grain boundaries from image data and shows the improvement of subsequent morphological filtering \citep{dramsch2018gaussian}. These methods reduce labor-intensive manual tasks, introducing varying degrees of automation in geoscience workflows.

The foundational work investigates low-frequency aliasing in \aclp{cnn} \citep{dramsch2018information} and goes on to show that phase information in complex-valued neural network can stabilize the reconstruction of compressed seismic data. The smaller complex-valued network outperforms larger real-valued networks, however, a very large real-valued network can implicitly learn partial phase information \citep{dramsch2019complex}. The paper touches on deficits of current metrics applied to geoscience and exposes a periodic dimming effect of frequencies from neural networks that should be further investigated, particularly in the context of aliasing. This paper led to the creation of the open source software package \texttt{keras complex} to enable complex-valued deep learning in \texttt{\acl{tf}} (Manual in \ref{section:keras-complex}). The code was available in an older framework \citep{trabelsi2017deep}, which I consolidated for this study. I went on to package the code, make it easily installable and generate online documentation to enable researchers to generate their own studies from this work \citep{dramsch2019complexcode}.

The research in \citet{dramsch2018deep} showed that transfer learning can alleviate the necessity for large amounts of labeled data, by re-using a neural network on natural images. This study showed the generalizable networks can be transferred to seismic data and outperform smaller networks trained from scratch, the smaller network size being necessary to avoid overfitting. The source code for this research was made available and has been of use to multiple researchers \citep{dramsch2018deepcode}. This has wide applications in industry and research settings, considering the limited availability of labeled data and wide availability of pre-trained network architectures. Moreover, this insight is applicable to pre-training geoscientific \aclp{nn} and fine-tuning these models to specific applications when needed.

The tutorial paper \citep{dramsch2019dtw} contributes insights into applications of \acl{dtw} in 4D seismic. This work explores the influence of varying metrics in geoscience, introducing the Huber loss as a possible loss function for geoscientific application. The study goes on to show that a novel application of the LB\_Keogh lower bound for \acl{dtw} can significantly improve the accuracy for field data, introducing a constraint on the search space for \acl{dtw}. The code for this tutorial including further interactive explorations into constraints for \acl{dtw} will be made available after the double-blind peer review process concludes \citep{dramsch2020dtwcode}.

The first application of \acl{ml} to 4D seismic data introduces a novel method to perform pressure-saturation inversion on amplitude difference maps \citep{dramsch2019deep}. This work introduces basic physics principles into the neural network architecture, which was shown to stabilize the training result. Moreover, this work shows the possibility of training \aclp{dnn} on simulation data and subsequently transferring the network to field data, by applying adequate noise injection. The \acl{dnn} results were successfully compared to results from the Bayesian inversion showing a promising application of \aclp{dnn} in 4D \acl{qi} \citep{dramsch2019deep}. While this work has attracted interest in a sponsors meeting and the workshop presentations \citep{dramsch2019deep,dramsch2019including}, further investigation into model explainability and lower complexity baseline models is necessary \citep[in preparation][]{corte2019exploring,dramsch2020physics}.

The second application of \acl{ml} to 4D seismic data presents a completely novel method for time-shift extraction \citep{dramsch20193dwarping}. This method combines recent advancements in diffeomorphic mapping, \acl{dl} and unsupervised learning to introduce a 3D time shift extraction method including uncertainty values, where 1D extraction is the standard. The method is shown to work on 3D seismic post-stack data with strongly differing acquisition parameters, without supplying any time shift information. After applying the method, the 3D seismic volumes are aligned well, with the diffeomorphic constraint performing well on seismic data. This work tests the trained network on two other 3D seismic volume pairs to test the generalization of the \acl{cnn} after training. The two test sets show that the trained model on a single 3D seismic volume pair transfers well to the same field with different acquisition parameters and even a different field with vastly different geological setting. The code is openly available to foster further development and investigation of the method \citep{dramsch20203dwarpingcode}.

Overall, this thesis shows the impact \acl{dl} can have in geoscience with two novel methods for 4D seismic analysis. It shows the impact of fundamental research of signal processing and information theory in \aclp{dnn} and \aclp{cnn}. Moreover, it shows applications of building systems in label sparse environments to leverage technological advancements and the value of including prior physical insights into \acl{ml} workflows. This thesis builds heavily on open source software and aims to return some of the effort.

\nocite{haris_iqbal_2018_2526396}